---
title: "Robust Experimental Design"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Robust Experimental Design}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  fig.align = "center"
)
```

## Overview

`robustExpDesign` designs experiments that will be combined with observational
or other external evidence. Because the external evidence may be misspecified,
the design balances statistical precision against worst-case bias exposure.

A typical workflow is:

1. specify a target sensitivity vector `omega`, or matrix `Omega`;
2. provide covariance and experimental-variance inputs;
3. specify costs and feasibility restrictions;
4. solve a proportional-regret design problem;
5. inspect experiment selection, estimator weights, sample allocation, and
   regret components.

The package implements methods from Epanomeritakis and Viviano (2026), *Learning
What to Learn: Experimental Design when Combining Experimental with
Observational Evidence*, manuscript dated August 4, 2026.

## Installation and optional solvers

`Matrix` and `rlang` are required dependencies. Optional features use:

- `quadprog` for the non-Gurobi experiment-design backend;
- `CVXR` for optimized moment/GMM designs;
- `ggplot2` for sweep plots;
- Gurobi for audience regret, positive `n_min`, direct experimental weights,
  and large mixed-integer designs.

```{r, eval = FALSE}
install.packages(c("quadprog", "CVXR", "ggplot2"))
library(robustExpDesign)
```

## Baseline proportional-regret design

Consider three parameters. `Sigma_obs` is the covariance matrix of the external
estimates, and `v2[j] / n_j` is the variance of experimental estimate `j` when
sample size `n_j` is assigned to that experiment.

```{r, eval = FALSE}
Sigma_obs <- matrix(
  c(
    0.08, 0.01, 0.00,
    0.01, 0.05, 0.01,
    0.00, 0.01, 0.09
  ),
  nrow = 3,
  byrow = TRUE
)

v2 <- c(direct = 1.00, income = 1.40, wage = 1.10)
omega <- c(direct = 0.40, income = 1.00, wage = -0.70)

fit <- solve_minimax_design(
  Sigma_obs = Sigma_obs,
  v2 = v2,
  n_total = 500,
  omega = omega,
  costs = c(1, 1, 1.5),
  bias_weights = c(1, 1, 1),
  h = 2,
  min_experiments = 1,
  solver = "quadprog"
)
```

The objective is the maximum of normalized variance and bias regret. The main
outputs are:

```{r, eval = FALSE}
fit$selected
fit$x_opt
fit$gamma_opt
fit$s_opt
fit$a_exp_opt
fit$a_obs_opt
fit$n_opt
fit$alpha_ratio
fit$beta_ratio
fit$regret
```

## Evaluating a supplied design

`evaluate_design()` computes allocation, variance, and bias for a selection and
weighting rule supplied by the user.

```{r, eval = FALSE}
evaluated <- evaluate_design(
  Sigma_obs = Sigma_obs,
  v2 = v2,
  n_total = 500,
  omega = omega,
  x = c(1, 1, 0),
  gamma = c(0.8, 0.6, 0),
  costs = c(1, 1, 1.5)
)

evaluated$n_opt
evaluated$alpha
evaluated$beta
```

## Weighted ambiguity sets

`bias_weights` implements coordinate-specific bounds

```text
|b_j| <= bias_weights[j] * B.
```

Zero weights mark coordinates treated as not misspecified.

```{r, eval = FALSE}
fit_weighted <- solve_minimax_design(
  Sigma_obs = Sigma_obs,
  v2 = v2,
  n_total = 500,
  omega = omega,
  costs = c(1, 1, 1.5),
  bias_weights = c(1.0, 0.5, 2.0),
  h = 2,
  min_experiments = 1,
  solver = "quadprog"
)
```

## Feasibility restrictions

The feasible menu can be restricted with experiment-count bounds, eligibility
indicators, explicit feasible sets, or linear constraints on the selection
vector.

```{r, eval = FALSE}
selection_constraints <- list(
  A = matrix(c(1, 1, 0), nrow = 1),
  sense = "<=",
  rhs = 1
)

fit_restricted <- solve_minimax_design(
  Sigma_obs = Sigma_obs,
  v2 = v2,
  n_total = 500,
  omega = omega,
  costs = c(1, 1, 1.5),
  h = 2,
  selection_constraints = selection_constraints,
  min_experiments = 1,
  solver = "quadprog"
)
```

Positive `n_min` imposes a minimum allocation on each selected experiment and
currently requires Gurobi. The floor is applied to candidate designs and oracle
benchmarks.

```{r, eval = FALSE}
fit_floor <- solve_minimax_design(
  Sigma_obs = Sigma_obs,
  v2 = v2,
  n_total = 500,
  omega = omega,
  costs = c(1, 1, 1.5),
  h = 2,
  min_experiments = 1,
  n_min = 50,
  solver = "gurobi"
)
```

## Budget sweeps

`sweep_n_total()` evaluates a grid of budgets and experiment-count limits.

```{r, eval = FALSE}
sweep <- sweep_n_total(
  Sigma_obs = Sigma_obs,
  v2 = v2,
  omega = omega,
  n_grid = c(200, 300, 400, 500, 750, 1000),
  h_values = c(1, 2),
  costs = c(1, 1, 1.5),
  experiment_names = c("Direct", "Income", "Wage"),
  include_variance_design = TRUE,
  make_plots = TRUE,
  solver = "quadprog"
)

head(sweep$data$arms)
head(sweep$data$regret)
sweep$plots$allocation
sweep$plots$regret
```

## Moment selection and GMM weights

For general moment-loading problems, provide a moment Jacobian `Lambda`, moment
covariance `Sigma`, and target sensitivity `omega` or `Omega`.

With `optimize_W = FALSE`, the function evaluates a supplied weighting matrix.

```{r, eval = FALSE}
Lambda <- matrix(
  c(
    1.0, 0.0,
    0.0, 1.0,
    1.0, 0.5,
    0.5, 1.0
  ),
  nrow = 4,
  byrow = TRUE
)

Sigma_mom <- diag(c(0.05, 0.06, 0.10, 0.12))
Omega <- matrix(c(0.4, 1.0), nrow = 1)

fixed_moments <- solve_moment_design(
  Lambda = Lambda,
  Sigma = Sigma_mom,
  Omega = Omega,
  W = diag(4),
  biased_moments = c(3, 4),
  norm = "l2",
  optimize_W = FALSE
)
```

With `optimize_W = TRUE`, `CVXR` optimizes the equivalent linear GMM estimator
over the supplied moment sets or masks.

```{r, eval = FALSE}
optimized_moments <- solve_moment_design(
  Lambda = Lambda,
  Sigma = Sigma_mom,
  Omega = Omega,
  biased_moments = c(3, 4),
  bias_weights = c(0, 0, 1, 1.5),
  norm = "linf",
  moment_sets = list(c(1, 2, 3), c(1, 2, 4), c(1, 2, 3, 4)),
  optimize_W = TRUE,
  cvxr_solver = "CLARABEL"
)
```

## Audience-regret designs

`solve_audience_regret_design()` solves the Appendix C.2 finite-grid objective.
The scalarization grid represents

```text
lambda = B^2 / (1 + B^2)
```

and must include 0 and 1. The current implementation requires Gurobi.

```{r, eval = FALSE}
fit_audience <- solve_audience_regret_design(
  Sigma_obs = Sigma_obs,
  v2 = v2,
  n_total = 500,
  omega = omega,
  costs = c(1, 1, 1.5),
  bias_weights = c(1, 1, 1),
  h = 2,
  min_experiments = 1
)

fit_audience$r_opt
fit_audience$risk_by_lambda
```