---
title: "Robust Experimental Design"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Robust Experimental Design}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  fig.align = "center"
)
```

## Introduction

The package `robustExpDesign` implements methods for designing experiments that will be combined with observational or external evidence. The external evidence may be biased, so the design problem balances statistical precision against sensitivity to misspecification.

The package is designed for applications where the researcher has a target estimand, an initial set of external estimates, and a menu of possible experiments, treatment arms, sites, moments, or data sources. The output describes which experiments to run, how much sample to allocate to each selected experiment, and how to combine experimental and external evidence in the estimator.

The typical workflow is:

1. specify the target sensitivity vector `omega`, or a target-sensitivity matrix `Omega`;
2. provide observational covariance information and experimental variance inputs;
3. specify costs, experiment-count limits, and feasibility restrictions;
4. solve the minimax proportional-regret design problem;
5. inspect selected experiments, estimator weights, sample allocation, and regret components.

## Reference

This package implements methods from:

Epanomeritakis, A., & Viviano, D. (2026). *Learning What to Learn: Experimental Design when Combining Experimental with Observational Evidence*. Manuscript, July 9, 2026.

## Installation

Install the development version from GitHub with:

```{r, eval = FALSE}
install.packages("remotes")
remotes::install_github("ostasovskyi/robustExpDesign")
```

Load the package with:

```{r, eval = FALSE}
library(robustExpDesign)
```

Core utilities use `Matrix`, `rlang`, and `utils`.

```{r, eval = FALSE}
install.packages(c("Matrix", "rlang"))
```

The non-Gurobi baseline solver uses `quadprog`.

```{r, eval = FALSE}
install.packages("quadprog")
```

The moment-selection and audience-regret solvers use `CVXR`.

```{r, eval = FALSE}
install.packages("CVXR")
```

Gurobi is optional for the mixed-integer backend. After installing and licensing the Gurobi Optimizer, install its R package from the Gurobi installation directory.

```{r, eval = FALSE}
library(gurobi)
```

## Example workflow

We first set up a small three-parameter problem. The target sensitivity vector is `omega`. The matrix `Sigma_obs` is the covariance matrix for the observational estimates. The vector `v2` contains experimental variance constants, so arm `j` has variance `v2[j] / n_j` when allocated sample size `n_j`.

```{r, eval = FALSE}
library(robustExpDesign)

Sigma_obs <- matrix(
  c(
    0.08, 0.01, 0.00,
    0.01, 0.05, 0.01,
    0.00, 0.01, 0.09
  ),
  nrow = 3,
  byrow = TRUE
)

v2 <- c(direct = 1.00, income = 1.40, wage = 1.10)
omega <- c(direct = 0.40, income = 1.00, wage = -0.70)
```

We now solve the minimax proportional-regret design problem. The researcher can select at most two experiments, must select at least one experiment, and faces a per-unit cost of 1.5 for the wage experiment.

```{r, eval = FALSE}
fit <- solve_minimax_design(
  Sigma_obs = Sigma_obs,
  v2 = v2,
  n_total = 500,
  omega = omega,
  costs = c(1, 1, 1.5),
  bias_weights = c(1, 1, 1),
  h = 2,
  min_experiments = 1,
  solver = "quadprog"
)
```

After fitting, we can inspect the selected design and estimator.

```{r, eval = FALSE}
fit
fit$selected
fit$x_opt
fit$gamma_opt
fit$n_opt
fit$alpha_ratio
fit$beta_ratio
fit$regret
```

The entry `x_opt` records which experiments are selected. The entry `gamma_opt` records shrinkage weights. The entry `n_opt` gives the cost-aware experimental sample allocation.

## Objective components

For a candidate design, the package evaluates

```text
max(variance regret, bias regret).
```

The variance-regret component compares the design's variance index with the smallest feasible variance index. The bias-regret component compares the design's worst-case bias exposure with the smallest feasible bias exposure.

```{r, eval = FALSE}
fit$alpha_opt
fit$beta_opt
fit$alpha_star
fit$beta_star
fit$alpha_ratio
fit$beta_ratio
fit$regret
```

A variance-optimal benchmark can be computed under the same feasibility restrictions.

```{r, eval = FALSE}
fit_var <- solve_variance_design(
  Sigma_obs = Sigma_obs,
  v2 = v2,
  n_total = 500,
  omega = omega,
  costs = c(1, 1, 1.5),
  h = 2,
  min_experiments = 1,
  solver = "quadprog"
)

fit_var$selected
fit_var$n_opt
fit_var$regret
```

## Weighted bias radii

The package supports weighted ambiguity sets for observational bias. Passing `bias_weights` implements coordinate-specific bounds of the form

```text
|b_j| <= bias_weights[j] * B.
```

The aliases `kappa` and `k` are also accepted. These aliases are useful when matching notation in scripts and equations.

```{r, eval = FALSE}
fit_weighted <- solve_minimax_design(
  Sigma_obs = Sigma_obs,
  v2 = v2,
  n_total = 500,
  omega = omega,
  costs = c(1, 1, 1.5),
  kappa = c(1.0, 0.5, 2.0),
  h = 2,
  min_experiments = 1,
  solver = "quadprog"
)

fit_weighted$beta_opt
fit_weighted$regret
```

## Feasibility restrictions

The feasible experiment set can be restricted with `h`, `min_experiments`, `x_max`, explicit `feasible_sets`, or linear `selection_constraints`.

```{r, eval = FALSE}
selection_constraints <- list(
  A = matrix(c(1, 1, 0), nrow = 1),
  sense = "<=",
  rhs = 1
)

fit_restricted <- solve_minimax_design(
  Sigma_obs = Sigma_obs,
  v2 = v2,
  n_total = 500,
  omega = omega,
  costs = c(1, 1, 1.5),
  h = 2,
  x_max = c(1, 1, 1),
  selection_constraints = selection_constraints,
  min_experiments = 1,
  solver = "quadprog"
)

fit_restricted$selected
fit_restricted$n_opt
```

## Budget sweeps

`sweep_n_total()` runs the design solvers over a grid of budgets and experiment-count limits. It returns arm-level allocation and shrinkage rows, one regret row for each solver run, and optional plots.

```{r, eval = FALSE}
sweep <- sweep_n_total(
  Sigma_obs = Sigma_obs,
  v2 = v2,
  omega = omega,
  n_grid = c(200, 300, 400, 500, 750, 1000),
  h_values = c(1, 2),
  costs = c(1, 1, 1.5),
  experiment_names = c("Direct", "Income", "Wage"),
  include_variance_design = TRUE,
  make_plots = TRUE,
  solver = "quadprog"
)

head(sweep$data$arms)
head(sweep$data$regret)
sweep$plots$allocation
sweep$plots$regret
```

## Moment selection and GMM weights

The moment-selection routines are for settings where the researcher wants to combine several moments, instruments, sites, outcomes, or data sources. Some moments may be reliable, while others may be useful but potentially biased.

The main function is `solve_moment_design()`. The user provides `Lambda`, the moment Jacobian; `Sigma`, the covariance matrix of the moments; and `omega` or `Omega`, which describes how the target estimand depends on structural parameters.

### Fixed weighting matrix

Use `solve_moment_design(..., optimize_W = FALSE)` to evaluate a supplied GMM weighting matrix.

```{r, eval = FALSE}
Lambda <- matrix(
  c(
    1.0, 0.0,
    0.0, 1.0,
    1.0, 0.5,
    0.5, 1.0
  ),
  nrow = 4,
  byrow = TRUE,
  dimnames = list(
    c("exp_direct", "exp_income", "obs_joint", "obs_aux"),
    c("theta_direct", "theta_income")
  )
)

Sigma_mom <- diag(c(0.05, 0.06, 0.10, 0.12))
Omega <- matrix(c(0.4, 1.0), nrow = 1)
W <- diag(4)

eval_fixed <- solve_moment_design(
  Lambda = Lambda,
  Sigma = Sigma_mom,
  Omega = Omega,
  W = W,
  biased_moments = c(3, 4),
  norm = "l2",
  optimize_W = FALSE
)

eval_fixed$alpha
eval_fixed$beta
```

### Optimizing over moment sets

Use `optimize_W = TRUE` to optimize over the first-order linear estimator implied by selected moments. Candidate moment sets or masks define the admissible moment menus.

```{r, eval = FALSE}
fit_moments <- solve_moment_design(
  Lambda = Lambda,
  Sigma = Sigma_mom,
  Omega = Omega,
  biased_moments = c(3, 4),
  bias_weights = c(0, 0, 1, 1.5),
  norm = "linf",
  moment_sets = list(c(1, 2, 3), c(1, 2, 4), c(1, 2, 3, 4)),
  optimize_W = TRUE,
  cvxr_solver = "CLARABEL"
)

fit_moments$selected_candidate
fit_moments$alpha_ratio
fit_moments$beta_ratio
fit_moments$regret
```

## Audience-regret designs

`solve_audience_design()` solves the audience-regret objective over a supplied grid of bias radii `B_grid`. The function computes oracle benchmarks at each grid value and then selects one experiment set and one sample allocation that control regret across the grid.

```{r, eval = FALSE}
fit_audience <- solve_audience_design(
  Sigma_obs = Sigma_obs,
  v2 = v2,
  n_total = 500,
  omega = omega,
  B_grid = c(0, 0.25, 0.5, 1, 2),
  costs = c(1, 1, 1.5),
  kappa = c(1, 1, 1),
  h = 2,
  min_experiments = 1,
  cvxr_solver = "CLARABEL"
)

fit_audience
fit_audience$n_opt
fit_audience$posterior_ratio
fit_audience$beta_ratio_large_B
```

For general linear reported statistics, pass `Lambda_obs` and `Lambda_exp`. The function then imposes the calibration constraint implied by those loading matrices.

## Function map

The main user-facing functions are:

```{r, eval = FALSE}
solve_minimax_design()
solve_variance_design()
evaluate_design()
sweep_n_total()
solve_moment_design()
solve_audience_design()
```

`solve_minimax_design()` is the main experiment-selection function. `solve_variance_design()` computes a variance-optimal benchmark. `evaluate_design()` evaluates a user-supplied design. `sweep_n_total()` runs budget and experiment-count sweeps. `solve_moment_design()` handles moment-based designs and GMM estimators. `solve_audience_design()` handles grid-based audience-regret designs.
